Constraints on Cosmic-Ray Particle Populations, Magnetic Fields, and Dark Matter from VHE Gamma-Ray Observations of the Coma Cluster of Galaxies with VERITAS

### 2011-07-21 ms.tex
Fixed a typo

\subsection{Simulation-based approach}
We adopt the universal spectral and spatial gamma-ray model developed by
\citet{article:PinzkePfrommer:2010} to estimate the emission from decaying neutral pions
that dominates over the IC emission from primary and secondary electrons above
100~MeV in clusters. Given a density profile as, e.g., inferred by cosmological
simulations or X-ray observations, the analytic approach models the CR
distribution and the associated radiative emission processes from radio to the
gamma-ray band.  This formalism was derived from high resolution simulations of
clusters of galaxies that included radiative hydrodynamics, star formation,
supernova feedback, and followed CR physics using a novel formulation that trace
the most important injection and loss processes self-consistently while
accounting for the CR pressure in the equation of motion
\citep{article:PfrommerSpringelEnsslinJubelgas, article:EnsslinPfrommerSpringelJubelgas:2007, article:JubelgasSpringelEnsslinPfrommer:2008}.  The
results are in line with earlier numerical results on some of the overall
characteristics of the CR distribution and the associated radiative emission
processes \citep{article:DolagEnsslin, article:MiniatiRyuKangJones:2001, article:Miniati:2003, article:Pfrommer_etal:2007, article:PfrommerEnsslinSpringel:2008, article:Pfrommer:2008}, but for the
first time is able to track the spectral evolution as a function of space while
accounting for the non-linear feedback of such a population on the hydrodynamics
in a high-resolution cosmological simulation of the formation of galaxy
clusters.

We note that the overall normalization of the CR and gamma-ray distribution
scales with the maximum acceleration efficiency at structure formation shock
waves. Following recent observations at supernova remnants
\cite{article:Helder_etal:2009} as well as theoretical studies
\cite{article:KandJones:2005}, we adopt an optimistic but nevertheless realistic
value of this parameter and assume that 50\% of the dissipated energy at strong
shocks is injected into CRs while this efficiency rapidly decreases for weaker
shocks \cite{article:EnsslinPfrommerSpringelJubelgas:2007}. Since the vast majority of internal formation
shocks (merger and flow shocks) are weak shocks with Mach numbers $M\lesssim3$
\citep[e.g.,][]{article:Ryu_etal:2003, article:PfrommerSpringelEnsslinJubelgas} they do not
significantly contribute to the CR population in clusters. Instead, strong
shocks during the formation epoch of clusters and strong accretion shocks at the
present time (at the boundary of voids and filaments/super-cluster regions)
dominate the injection of CRs which are adiabatically transported into clusters
thereafter. Hence, these model predictions provide a plausible upper limit for
the CR contribution from structure formation shocks in galaxy clusters which can
be scaled down for a lower effective acceleration efficiency in the strong shock
regime.  On the other hand, other possible CR sources such as AGN and starburst
driven galactic winds have been neglected for simplicity which could in
principle increase the expected gamma-ray yield from clusters by an unknown
factor.

There are additional limitations of the simulated physics in these cosmological
simulations which only considered advective transport of CRs by turbulent gas
motions which produces centrally enhanced profiles.  However, active CR
transport such as CR diffusion and streaming tends to drive the CR radial
profiles towards being flat, with equal CR number density everywhere. While the
CR streaming is then usually larger than typical advection velocities and
becomes comparable or lower than this only for periods with trans- and
supersonic cluster turbulence during a cluster merger. As a consequence a
bimodality of the CR spatial distribution results with merging (relaxed)
clusters showing a centrally concentrated (flat) CR energy density profile
\cite{article:EnsslinPfrommerMiniatiSubramanian:2011}.  This translates into a bimodality of the expected
diffuse radio and gamma-ray emission of clusters, since more centrally
concentrated CR will find higher target densities for hadronic CR proton
interactions \cite{article:EnsslinPfrommerMiniatiSubramanian:2011}. As a result of this, relaxed clusters
could have a reduced gamma-ray luminosity by up to a factor of five. The goal of
this and other future work of seeking for gamma-ray emission from clusters is to
constrain the acceleration physics and transport properties of CRs.

We adopt the electron density profile for the Coma cluster that has been
inferred from ROSAT X-ray observations \citep{article:BrielHenryBohringer:1992} and model the
spatial and spectral distribution of CRs according to the (conservative) model
by \citet{article:PinzkePfrommer:2010} that neglects the contribution of cluster
galaxies. Figure ?? shows the expected spectrum of Coma, normalized to the total
flux within the virial radius. Add {\em Fermi} results \ldots

Figure~\ref{fig:xcr} shows the relative CR pressure, $X_\CR = \bra P_\CR
\ket / \bra P_\rmn{th} \ket$ as a function of radial distance from the
Coma cluster center (dashed lines). To compute the CR pressure, we assume a
lower cutoff of the CR distribution of $P = 0.8\,m_\rmn{p} c^2$ as found in
these simulations due to the high cooling rates at low CR energies that
effectively thermalizes the low-energy, trans-relativistic CR population.  The
relative CR pressure rises toward the outer regions due to CR acceleration being
more efficient at the peripheral strong accretion shocks compared to weak
central flow shocks and adiabatic compression of a composite of CRs and thermal
gas disfavours the CR pressure relative to the thermal pressure due to the
softer equation of state of CRs. The weak increase of $X_\CR$ toward the core is
due to the comparably fast thermal cooling of gas which diminishes the thermal
pressure support relative to that in CRs yielding to adiabatic compression of
CRs.

Table~\ref{table:constraints} shows the expected gamma-ray fluxes within the
regions of interest. For the most constraining regions tested (within an
aperture of radius 0.2$^\circ$), those are a factor of 7.2 below the inferred
upper limits (see Table~\ref{table:results} and assuming a spectral index of
$\alpha=2.3$ which matches the simulated one at energies $E=200$~GeV).  Hence we
scale the averaged relative CR pressure, $X_\CR = \bra P_\CR \ket /
\bra P_\rmn{th} \ket$ of our model by that factor, keep the spatial
behavior and obtain an integrated limit of the relative CR pressure of
$X_\CR<0.091$ within 0.2$^\circ$ that translates to a limit within the cluster
virial radius of $X_\CR<0.13$ (solid lines of Figure~\ref{fig:xcr}).

\subsection{Simplified analytical model}

To complement the simulation-inspired analysis and to facilitate comparison with
earlier analytical work, we repeat the data analysis with a simplified
analytical model that assumes a power-law CR spectrum and a constant relative CR
pressure, i.e., we adopt the {\em isobaric model of CRs}. Since a priori, the CR
spectral index is unconstrained, we vary it within plausible ranges of
$2.1<\alpha_\rmn{p}<2.5$. These spectral ranges are needed to explain the radio
spectral index of the giant radio halo of Coma after accounting for the spectral
steepening at frequencies $\nu\sim5~\rmn{GHz}$ due to the Sunyaev-Zel'dovich
effect \citep{article:PfrommerEnsslin:2004b}\footnote[4]{Assuming a magnetic field of
  1~$\mu$G, the CR protons responsible for the GHz radio emitting electrons have
  an energy of $~100$~GeV and are $\sim$ 20 times less energetic than those CR
  protons that are responsible for the 200 GeV $\gamma$-ray emission. This is
  consistent with the concave curvature found in the universal CR spectrum by
  \citet{article:PinzkePfrommer:2010}.}.  Table~\ref{table:constraints} shows the
resulting constraints on $X_\CR$ which range between 0.05 and 1, strongly
depending on the assumed spectral index.

Add {\em Fermi} results, critical spectral index \ldots

%the limit for 0.2 deg and alpha_p = 2.3 is compatible with the simulation based
%one, X_CR < 0.1. However, there is currently a discrepancy in our assumptions
%about the lower energy cutoff of CRs: in the simulations we use P/mp c = 0.8 (as
%simulated) while I set this to 0 in my simplified (isobaric) model. The
%discrepancy can be understood as follows:
%*) in the isobaric model, we would have to normalize our $X_\CR$ by a factor of 2/3
%for an effective spectral index of ~ 2.4 (between GeV and 200 GeV) as inferred
%from the simulations \citep[see Fig 1 in][]{article:PfrommerEnsslin:2004}).
%*) In the simulation model, we have an increased pressure by this factor of 3/2
%(in comparison to a power law spectrum with alpha_TeV ~ 2.3 down to a cutoff in
%the normalized momentum of 0.8) due to the concave curvature of our spectrum.
%}

\subsection{Minimum gamma-ray flux}

For clusters that host radio halos we are able to derive a minimum $\gamma$-ray
flux in the hadronic model of CR interactions. The idea is based on the fact
that a stationary distribution of CR electrons loses all its energy to
synchrotron radiation for strong magnetic fields ($B \gg B_\rmn{CMB} \simeq 3.2
\mu$G) so that the ratio of $\gamma$-ray to synchrotron flux becomes independent
of the spatial distribution of CRs and thermal gas \citep{article:Pfrommer:2008}.
This can be easily seen by considering the pion decay induced $\gamma$-ray
luminosity $L_\gamma$ and the synchrotron luminosity $L_\nu$ of a steady state
distribution of CR electrons that has been generated by hadronic CR
interactions,
\begin{eqnarray}
L_\gamma &=& A_\gamma \dps\int d V\, n_\CR n_\rmn{gas},\\
\label{eq:Lnu}
   L_\nu &=& A_\nu \dps\int d V\, n_\CR n_\rmn{gas} \frac{\dps
    \eps_B^{(\alpha_\nu+1)/2}}{\dps\eps_\rmn{CMB}+\eps_B}\\
  &\simeq&A_\nu \dps\int d V\, n_\CR n_\rmn{gas}\quad\rmn{for}~\eps_B \gg \eps_\rmn{CMB}.
\end{eqnarray}
Here $A_\gamma$ and $A_\nu$ are dimensional constants that depend on the
hadronic physics of the interaction \citep{article:PfrommerEnsslinSpringel:2008,
  article:Pfrommer:2008} and $\alpha_\nu\simeq 1$ is the observed synchrotron
spectral index. Hence we can derive a minimum $\gamma$-ray flux in the hadronic
model
\begin{equation}
  \label{eq:Fmin}
  \mathcal{F}_{\gamma,\rmn{min}} = 
  \frac{\dps A_\gamma}{\dps A_\nu}\frac{\dps L_\nu}{\dps 4\pi D_\rmn{lum}^2},
\end{equation}
where $L_\nu$ is the observed luminosity of the radio mini-halo and
$D_\rmn{lum}$ denotes the luminosity distance to the respective cluster.
Lowering the magnetic field would require an increase in the energy density of
CR electrons to reproduce the observed synchrotron luminosity and thus increase
the associated $\gamma$-ray flux.

To derive a minimum gamma-ray flux that is comparable to the upper limits, we
need to determined the radio flux within the corresponding angular regions. To
this end, we fit the point source subtracted, azimuthally averaged radio halo
profile at 1.38 GHz \citep{article:Deiss_etal:1997} with a $\beta$-model,
\begin{equation}
S_\nu (r_\bot)= S_0\, 
\left[ 1 + \left( \frac{r_\bot}{r_{\rmn{c}}}\right)^2\right]
^{-3\beta + 1/2},
\label{beta}
\end{equation}
where $S_0 = 1.1 \times 10^{-3}\,\rmn{Jy\,arcmin}^{-2}$, $r_{\rmn{c}} = 450
kpc$, and $\beta = 0.78$.  Within the error bars, this profile is consistent
with 326 MHz-data taken by \citet{article:Govoni_etal:2001} when scaled with a radio
spectral index of 1.15.

The results for the minimum gamma-ray flux $F_{\rmn{min}}(>E=220~\rmn{GeV}$ and
the minimum relative CR pressure, $X_{\CR,\,\rmn{min}} = F_{\rmn{min}}/F$ are
shown in Table~\ref{table:constraints}. Even in the most constraining case
($\alpha=2.1$, $<0.2^\circ$), these are a factor of about 60 below the upper
limits obtained in this work.

Add {\em Fermi} results\ldots

\section{Constraining the Magnetic Field}
\label{sec:B}

We have seen in the previous Section that we can obtain an absolute lower limit
on the gamma-ray emission in the hadronic model by assuming high magnetic
fields, $B>B_\rmn{CMB}$. Hence, we can turn this around by using our upper limit
on the gamma-ray emission (and by extension on the CR pressure) to infer a lower
limit on the magnetic field that is needed to explain the observed radio
emission which, however, assumes the validity of the hadronic model. Lowering
the gamma-ray constraint will tighten (increase) the magnetic field limit. Once
this gets in conflict with magnetic field measurements by means of other
methods, e.g., Faraday rotation measure (RM)\footnote{Generally, Faraday RM
  of the magnetic field strength by, e.g., background sources through clusters
  are degenerate with the magnetic coherence scale. Only a simultaneous
  deprojection method that self-consistently solves for the magnetic power
  spectrum is in principle able to break this degeneracy, however, require
  high-quality data and picks up further complications when using embedded
  sources within the cluster plasma, e.g., degeneracies with the assumed
  projection angle of the jet \citep{article:KucharEnsslin:2011}.}, this would
challenge the hadronic model of radio halos.  The method that we use to
constrain the magnetic field inherits a dependence on the assumed spatial
structure that we parametrize by
\begin{equation}
  \label{eq:B}
  B(r) = B_0  \,\left(\frac{n_\rmn{e}(r)}{n_\rmn{e}(0)}\right)^{\alpha_B},
\end{equation}
as suggested by Faraday RM studies and numerical magneto-hydrodynamical (MHD)
simulations \citep[][and references therein]{article:Bonafede_etal:2010, article:Bonafede_etal:2011}. In fact, the magnetic field in the Coma cluster is among
the best ones constrained due to its proximity that enables RM observations of
seven radio-sources located at projected distances of 50 to 1500 kpc from the
cluster center. The best fit model obtains $B_0 = 4.7^{+0.7}_{-0.8}\,\mu$G and
$\alpha_B = 0.5^{+0.2}_{-0.1}$ \citep{article:Bonafede_etal:2010}.  We aim for
constraining the central field strength, $B_0$, which depends on the magnetic
decline, $\alpha_B$, that we vary (within reasonable ranges of
$\Delta\alpha_B=0.2$ as suggested from the Faraday RM studies) to address this
dependence. We proceed as follows:

1) Given a model for the magnetic field, we determine the profile of the
relative CR pressure, $X_\CR$, so that the hadronically produced CR electrons
exactly matches the observed radio halo emission over the entire extent. To this
end, we de-project the fit to the surface brightness profile (using an Abel
integral equation, see Appendix of \citealt{article:PfrommerEnsslin:2004b}) yielding the
radio emissivity,
\begin{equation}
  \label{eq:Coma:radio}
  j_\nu (r) = \frac{S_0}{2\pi\, r_\rmn{c}}\, 
  \frac{6\beta - 1}{\left(1 + r^2/r_\rmn{c}^2\right)^{3 \beta}}\,
  \mathcal{B}\left(\frac{1}{2}, 3\beta\right)
  = j_{\nu,0} \left(1 + r^2/r_\rmn{c}^2\right)^{-3 \beta},
\end{equation}
where $\mathcal{B}$ denotes the beta function. Using ``isobaric CR model'' of
the formalism suggested by \citet{article:PfrommerEnsslin:2004b} to compute the secondary
radio emission and imposing the boundary condition to reproduce the observed
radio emission profile uniquely determines $X_\CR(r)$. One of those realizations
for the relative CR energy density $\tilde{X}_\CR(r) = \eps_\CR (r) /
\eps_\rmn{th} (r)$ and $\tilde{X}_B(r) = \eps_B (r) / \eps_\rmn{th} (r)$ is
shown in Figure 2 (lower left panel) of \citet{article:PfrommerEnsslin:2004a} which also
has the interesting property to represent the minimum energy configuration that
is able to reproduce the radio emission in the hadronic model.\footnote{Note
  that a central magnetic field of $3\,\mu$G corresponds in the Coma cluster to
  a relative magnetic energy density of $\tilde{X}_B=0.005$. For a CR population
  in clusters that have been accelerated in structure formation shocks the
  relativistic limit $E_\CR/E_\rmn{th} = 2 \bra P_\CR \ket / \bra P_\rmn{th}
  \ket$ is not applicable since the CR pressure is dominated by the
  trans-relativistic regime. This implies a somewhat harder equation of state
  for the CRs with a larger adiabatic index and yields the relation
  $E_\CR/E_\rmn{th} = 1.6 \bra P_\CR \ket / \bra P_\rmn{th} \ket$.}  The shaded
regions indicate the logarithmic curvature at the minimum, showing that only
modest additional energy investment would be needed to move the system by a
factor two away from this energetically least expensive state which needs not to
be realized in Nature.  The realization of the minimum energy magnetic field
corresponds to a magnetic decline of $\alpha_B\simeq0.3$ and is close to
equipartition with the CR proton pressure (but not that of the resulting
secondary CR electrons).  It is generically true for weak magnetic fields
($B<B_\rmn{CMB}$) in the outer parts of the Coma halo that the product
$X_\CR(r)X_B(r)$ has to increase by a factor of about 100 toward the radio halo
periphery to account for the observed extent.  If we were to adopt a steeper
magnetic decline such as $\alpha_B=0.5$ which produces a flat $X_B(r)$, the
relative CR pressure would have to rise accordingly by a factor of 100.

2) Given this realization for $X_\CR$, we compute the pion decay gamma-ray
surface brightness profile, integrate the flux within a radius of $\{0.13, 0.2,
0.4\}$ degree, and scale the CR profile in order to match the corresponding
VERITAS upper limits.  This scaling factor, $X_{\CR,0}$, depends on the CR
spectral index, $\alpha$, (assuming a power-law CR population for simplicity) and
the radial decline of the magnetic field, $\alpha_B$.

3) In the last step, we scale the secondary radio emission profile linearly
with $X_{\CR,0}$ and solve for $B_0$ subject to the boundary condition to match
the observed synchrotron profile. Note that for $B_0 \gg B_\rmn{CMB}$ and radio
spectral index, $\alpha_\nu$, the solution would be degenerate according to
Eqn.~(\ref{eq:Lnu}).

>>>>

\subsection{Simulation-based approach}
We adopt the universal spectral and spatial gamma-ray model developed by \citet{article:PinzkePfrommer:2010} to estimate the emission from decaying neutral pions that dominates over the IC emission from primary and secondary electrons above 100 MeV in clusters. Given a density profile as, e.g., inferred by cosmological simulations or X-ray observations, the analytic approach models the CR distribution and the associated radiative emission processes from radio to the gamma-ray band. This formalism was derived from high resolution simulations of clusters of galaxies that included radiative hydrodynamics, star formation, supernova feedback, and followed CR physics by tracing the most important injection and loss processes self-consistently while accounting for the CR pressure in the equation of motion \citep{article:PfrommerSpringelEnsslinJubelgas, article:EnsslinPfrommerSpringelJubelgas:2007, article:JubelgasSpringelEnsslinPfrommer:2008}. The results are in line with earlier numerical results on some of the overall characteristics of the CR distribution and the associated radiative emission processes \citep{article:DolagEnsslin, article:MiniatiRyuKangJones:2001, article:Miniati:2003, article:Pfrommer_etal:2007, article:PfrommerEnsslinSpringel:2008, article:Pfrommer:2008}.

We note that the overall normalization of the CR and gamma-ray distribution scales with the maximum acceleration efficiency at structure formation shock waves. Following recent observations at supernova remnants \citep{article:Helder_etal:2009} as well as theoretical studies \citep{article:KangJones:2005}, we adopt an optimistic but nevertheless realistic value of this parameter and assume that 50\% of the dissipated energy at strong shocks is injected into CRs while this efficiency rapidly decreases for weaker shocks. Since the vast majority of internal formation shocks (merger and flow shocks) are weak shocks with Mach numbers $M\lesssim3$ \citep[e.g.,][]{article:Ryu_etal:2003}, they do not significantly contribute to the CR population in clusters. Instead, strong shocks during the formation epoch of clusters and strong accretion shocks at the present time (at the boundary of voids and filaments/super-cluster regions) dominate the injection of CRs which are adiabatically transported through the cluster. Hence, these model predictions provide a plausible upper limit for the CR contribution from structure formation shocks in galaxy clusters which can be scaled with the effective acceleration efficiency. Other possible CR sources, such as AGN and starburst-driven galactic winds have been neglected for simplicity which could in principle increase the expected gamma-ray yield.

There cosmological simulations only considered advective transport of CRs by turbulent gas motions which produces centrally enhanced profiles.  However, other means of CR transport such as diffusion and streaming may flatten the CR radial profiles. It has been suggested that advection is dominant only during mergers \citep{article:EnsslinPfrommerMiniatiSubramanian:2011}, in which case relaxed clusters could have a reduced gamma-ray luminosity by up to a factor of five compared with merging clusters. The goal of this and other future work of seeking for gamma-ray emission from clusters is to constrain the acceleration physics and transport properties of CRs.

We adopt the electron density profile for the Coma cluster that has been inferred from ROSAT X-ray observations \citep{article:BrielHenryBohringer:1992} and model the spatial and spectral distribution of CRs according to the model by \citet{article:PinzkePfrommer:2010} that neglects the contribution of cluster galaxies. Figure ?? shows the expected spectrum of Coma, normalized to the total flux within the virial radius. [Add {\em Fermi} results]

Figure \ref{fig:xcr} shows the relative CR pressure, $X_{\CR} = \expval{P_{\CR}}/\expval{P_\rmn{th}}$ as a function of radial distance from the Coma cluster center (dashed lines). To compute the CR pressure, we assume a low-energy cutoff of the CR distribution at $T = 0.8\,m_\rmn{p} c^2$ reflecting the high cooling rates at low CR energies. The relative CR pressure rises toward the outer regions on account of the higher efficiency of CR acceleration at the peripheral strong accretion shocks compared to weak central flow shocks. Adiabatic compression of a mixture of CRs and thermal gas disfavours the CR pressure relative to the thermal pressure on account of the softer equation of state of CRs. The weak increase of $X_{\CR}$ toward the core is due to the comparably fast thermal cooling of gas which diminishes the thermal support relative to that in CRs.

Table \ref{table:constraints} shows the expected gamma-ray fluxes within the regions of interest. For the most constraining regions tested (within an aperture of radius 0.2$^{\circ}$), those are a factor of 7.2 below the inferred upper limits (see Table \ref{table:results} and assuming a spectral index of $\alpha=2.3$ which matches the simulated one at energies $E=200$ GeV). Hence we scale the averaged relative CR pressure, $X_{\CR}$
% = \bra P_\CR \ket /\bra P_\rmn{th} \ket$ 
of our model by that factor, keep the spatial behavior and obtain an integrated limit of the relative CR pressure of $X_{\CR}<0.091$ within 0.2$^{\circ}$ that translates to a limit within the cluster virial radius of $X_\CR<0.13$ (solid lines of Figure \ref{fig:xcr}).

\subsection{Simplified analytical model}
To complement the simulation-inspired analysis and to facilitate comparison with earlier analytical work, we repeat the data analysis with a simplified analytical model that assumes a power-law CR spectrum and a constant relative CR pressure, i.e., we adopt the {\em isobaric model of CRs}. Since a priori, the CR spectral index is unconstrained, we vary it within plausible ranges of $2.1<\alpha_{\rmn{p}}<2.5$. This spectral range is compatible with the radio spectral index of the giant radio halo of Coma after accounting for the spectral
steepening at frequencies $\nu\sim5~\rmn{GHz}$ due to the Sunyaev-Zel'dovich effect \citep{article:PfrommerEnsslin:2004b}\footnote{Assuming a magnetic field of 1 $\mu$G, the CR protons responsible for the GHz radio emitting electrons have an energy of $\sim100$ GeV and are $\sim$ 20 times less energetic than those CR protons responsible for 200-GeV gamma-ray emission.}. Table \ref{table:constraints} shows the resulting constraints on $X_\CR$ which range between 0.05 and 1, strongly
depending on the assumed spectral index.

[Add {\em Fermi} results, critical spectral index]

%
% This part was in \del{...} 
%
%the limit for 0.2 deg and alpha_p = 2.3 is compatible with the simulation based
%one, X_CR < 0.1. However, there is currently a discrepancy in our assumptions
%about the lower energy cutoff of CRs: in the simulations we use P/mp c = 0.8 (as
%simulated) while I set this to 0 in my simplified (isobaric) model. The
%discrepancy can be understood as follows:
%*) in the isobaric model, we would have to normalize our $X_\CR$ by a factor of 2/3
%for an effective spectral index of ~ 2.4 (between GeV and 200 GeV) as inferred
%from the simulations \citep[see Fig 1 in][]{article:PfrommerEnsslin:2004}).
%*) In the simulation model, we have an increased pressure by this factor of 3/2
%(in comparison to a power law spectrum with alpha_TeV ~ 2.3 down to a cutoff in
%the normalized momentum of 0.8) due to the concave curvature of our spectrum.
%}

\subsection{Minimum gamma-ray flux}
For clusters that host radio halos we are able to derive a minimum gamma-ray flux in the hadronic model of CR interactions. The idea is based on the fact that a stationary distribution of CR electrons loses all its energy to synchrotron radiation for strong magnetic fields ($B \gg B_{\rmn{CMB}} \simeq 3.2 \mu$G) so that the ratio of gamma-ray to synchrotron flux becomes independent of the spatial distribution of CRs and thermal gas \citep{article:Voelk:1989, article:Pohl:1994}, in particular with $\alpha_{\nu}\simeq 1$ as the observed synchrotron spectral index.  Hence we can derive a minimum gamma-ray flux in the hadronic
model
\begin{equation}
\label{eq:Fmin}
\mathcal{F}_{\gamma,\rmn{min}} = \frac{\dps A_{\gamma}}{\dps A_{\nu}}\frac{\dps L_{\nu}}{\dps 4\pi D_{\rmn{lum}}^{2}},
\end{equation}
where $L_{\nu}$ is the observed luminosity of the radio mini-halo and $D_{\rmn{lum}}$ denotes the luminosity distance to the respective cluster. Lowering the magnetic field would require an increase in the energy density of CR electrons to reproduce the observed synchrotron luminosity and thus increase the associated gamma-ray flux.

To derive a minimum gamma-ray flux that is comparable to the upper limits, we need to determined the radio flux within the corresponding angular regions. To this end, we fit the point-source subtracted, azimuthally averaged radio halo profile at 1.38 GHz \citep{article:Deiss_etal:1997} with a $\beta$-model,
\begin{equation}
\label{beta}
 S_{\nu} (r_{\bot})= S_{0} \left[ 1 + \left( \frac{r_{\bot}}{r_{\rmn{c}}}\right)^{2}\right]^{-3\beta + 1/2},
\end{equation}
where $S_{0} = 1.1 \times 10^{-3}\,\rmn{Jy\,arcmin}^{-2}$, $r_{\rmn{c}} = 450$ kpc, and $\beta = 0.78$.  Within the error bars, this profile is consistent with 326-MHz data taken by \citet{article:Govoni_etal:2001} when scaled with a radio spectral index of 1.15.

The results for the minimum gamma-ray flux $F_{\rmn{min}}(>220~\rmn{GeV}$ and the minimum relative CR pressure, $X_{\CR,\,\rmn{min}} = F_{\rmn{min}}/F$ are shown in Table \ref{table:constraints}. Even in the most constraining case ($\alpha=2.1$, $<0.2^{\circ}$), these are a factor of about 60 below the upper limits obtained in this work.

[Add {\em Fermi} results]

\section{Constraining the Magnetic Field}
\label{sec:B}
We have seen in the previous Section that we can obtain an absolute lower limit on the gamma-ray emission in the hadronic model by assuming high magnetic fields, $B>B_\rmn{CMB}$. Hence, we can turn this around by using our upper limit on the gamma-ray emission (and by extension on the CR pressure) to infer a lower limit on the magnetic field that is needed to explain the observed radio emission which, however, assumes the validity of the hadronic model. Lowering the gamma-ray constraint will tighten (increase) the magnetic field limit. Once this gets in conflict with magnetic field measurements by means of other methods, e.g., Faraday rotation measure (RM)\footnote{Generally, Faraday RM analysis of the magnetic-field strength by, e.g., background sources through clusters are degenerate with the magnetic coherence scale and may be biased by the unknown correlation between magnetic and density fluctuations.}, this would challenge the hadronic model of radio halos. The method that we use to constrain the magnetic field inherits a dependence on the assumed spatial structure that we parametrize by
\begin{equation}
\label{eq:B}
B(r) = B_{0} \,\left(\frac{n_{\rmn{e}(r)}}{n_{\rmn{e}(0)}}\right)^{\alpha_B},
\end{equation}
as suggested by Faraday RM studies and numerical magneto-hydrodynamical (MHD) simulations \citep[][and references therein]{article:Bonafede_etal:2010, article:Bonafede_etal:2011}. In fact, the magnetic field in the Coma cluster is among the best constrained, because its proximity permits RM observations of
seven radio-sources located at projected distances of 50 to 1500 kpc from the cluster center. The best-fit model obtains $B_{0} = 4.7^{+0.7}_{-0.8}\,\mu$G and $\alpha_{B} = 0.5^{+0.2}_{-0.1}$ \citep{article:Bonafede_etal:2010}. We aim for constraining the central field strength, $B_{0}$, which depends on the magnetic decline, $\alpha_{B}$, that we vary (within reasonable ranges of $\Delta\alpha_{B}=0.2$ as suggested from the Faraday RM studies) to address this dependence. We proceed as follows:

\begin{enumerate}
\item Given a model for the magnetic field, we determine the profile of the relative CR pressure, $X_{\CR}$, so that the hadronically produced CR electrons exactly matches the observed radio halo emission over the entire extent. To this end, we de-project the fit to the surface brightness profile (using an Abel integral equation, see Appendix of \citealt{article:PfrommerEnsslin:2004b}) yielding the radio emissivity,
\begin{equation}
\label{eq:Coma:radio}
j_{\nu} (r) = \frac{S_{0}}{2\pi\, r_{\rmn{c}}}\,
\frac{6\beta - 1}{\left(1 + r^{2}/r_{\rmn{c}}^{2}\right)^{3 \beta}}\,
\mathcal{B}\left(\frac{1}{2}, 3\beta\right)
= j_{\nu,0} \left(1 + r^2/r_{\rmn{c}}^{2}\right)^{-3 \beta},
\end{equation}
where $\mathcal{B}$ denotes the beta function. Using ``isobaric CR model'' to compute the secondary radio emission profile uniquely determines $X_\CR(r)$. It is generically true for weak magnetic fields ($B<B_{\rmn{CMB}}$) in the outer parts of the Coma halo that the product $X_{\CR}(r)X_{B}(r)$ has to increase by a factor of about 100 toward the radio halo periphery to account for the observed extent. If we were to adopt a steeper magnetic decline such as $\alpha_{B}=0.5$ which produces a flat $X_{B}(r)$, the relative CR pressure would have to rise accordingly by a factor of 100.

\item Given this realization for $X_{\CR}$, we compute the pion-decay gamma-ray surface brightness profile, integrate the flux within a radius of $\{0.13, 0.2, 0.4\}$ degree, and scale the CR profile in order to match the corresponding VERITAS upper limits. This scaling factor, $X_{\CR,0}$, depends on the CR spectral index, $\alpha$, (assuming a power-law CR population for simplicity) and the radial decline of the magnetic field, $\alpha_{B}$.

\item In the last step, we scale the secondary radio emission profile linearly with $X_{\CR,0}$ and solve for $B_{0}$ subject to the boundary condition to match the observed synchrotron profile. Note that for $B_{0} \gg B_{\rmn{CMB}}$ and radio spectral index, $\alpha_{\nu}$, the solution would be degenerate according to Eqn. (\ref{eq:Lnu}). 
\end{enumerate}
